Kalman Filters: Derivation

### 2015-07-27 ms-thesis/ver2/kalman/ms-thesis-kalman.tex
-

\begin{figure}[!hbp]
\center{
\begin{tabbing}
{\tt LDS}\={\tt Inf}\={\tt erence}$(Y,A,C,Q,R,x,P)$\\
\> {\tt for } $t = 1$ {\tt to } $T$ \% forward recursion \\
\> \> $x_t^{t-1} = Ax_{t-1}^{t-1}$\\
\> \> $P_t^{t-1} = AP_{t-1}^{t-1}A' + Q$ {\tt if } $t>1$\\
\> \> $K_t = P_t^{t-1}C'(CP_t^{t-1}C'+R)^{-1}$\\
\> \> $x_t^t = x_t^{t-1} + K_t(y_t - Cx_t^{t-1})$\\
\> \> $P_t^t = P_t^{t-1} - K_tCP_t^{t-1}$\\
\> {\tt end}\\
\> {\tt initialize } $\hat{P}_{T,T-1}=(I-K_TC)AP_{T-1}^{T-1}$\\
\> \> {\tt for } $t=T$ {\tt to } $2$ \% backward recursion \\
\> \> $J_{t-1} = P_{t-1}^{t-1}A'(P_t^{t-1})^{-1}$\\
\> \> $\hat{x}_{t-1} = x_{t-1}^{t-1} + J_{t-1}(\hat{x}_{t} - Ax_{t-1}^{t-1})$\\
\> \> $\hat{P}_{t-1} = P_{t-1}^{t-1} + J_{t-1}(\hat{P}_t - P_t^{t-1})J_{t-1}'$\\
\> \> $\hat{P}_{t,t-1} = P_t^t J_{t-1}' + J_t(\hat{P}_{t+1,t} - AV_t^t)J_{t-1}'$
   {\tt if } $t<T$\\
\> {\tt end}\\
\> {\tt return }$\hat{x}_t, \hat{P}_t, \hat{P}_{t,t-1}$ {\tt for all } $t$
\end{tabbing}
}
\caption{LDS Inference Algorithm}
\vspace{0.6cm}
\end{figure}

\begin{figure}[!hbp]
\center{
\begin{tabbing}
{\tt LDS}\={\tt Learn}\=$(Y,k,$\=$\epsilon)$\\ \> {\tt initialize } $A, C, Q, R,
x_1^0, P_1^0$\\ \> $\alpha = \sum_t y_ty_t'$\\ \> {\tt while change in
  likelihood >} $\epsilon$\\ \> \> {\tt LDSInference}$(Y,A,C,Q,R,x_1^0,P_1^0)$
\% E-step\\ \> \> {\tt initialize } $\delta = 0, \gamma = 0, \sigma = 0$\\ \> \>
   {\tt for } $t = 1$ {\tt to } $T$\\ \> \> \> $\delta = \delta +
   y_t\hat{x}_t$\\ \> \> \> $\gamma = \gamma + \hat{x}_t\hat{x}_t' +
   \hat{P}_t$\\ \> \> \> $\beta = \beta + \hat{x}_t\hat{x}_{t-1}' + P_{t,t-1}$
       {\tt if } $t>1$\\ \> \> {\tt end}\\ \> \> $\gamma_1 = \gamma -
       \hat{x}_T\hat{x}_T' - \hat{P}_T$\\ \> \> $\gamma_2 = \gamma -
       \hat{x}_1\hat{x}_1' - \hat{P}_1$\\ \> \> \% M-step\\ \> \> $C = \delta
       \gamma^{-1}$\\ \> \> $R = (\gamma - C\delta')/T$\\ \> \> $A =
       \beta\gamma_1^{-1}$\\ \> \> $Q = (\gamma_2 - A\beta') / (T-1)$\\ \> \>
       $x_1^0 = \hat{x}_1$\\ \> \> $P_1^0 = \hat{P}_1$\\ \> {\tt end}\\ \> {\tt
         return } $A, C, Q, R, x_1^0, P_1^0$
\end{tabbing}
}
\caption{LDS Learning Algorithm}
\vspace{0.6cm}
\end{figure}

@@

\begin{figure}[!hbp]
\center{
\begin{tabbing}
{\tt ob}\={\tt ser}\={\tt vation}\={\tt Genera}\={\tt tion}$(A,C,R,Q,\mu,\Sigma,T,N)$\\
\> $mu_x = mu[T-1];$\\
\> $sigma_x = sigma[T-1];$\\
\> {\tt for} $i = 1$ {\tt to } $N$\\
\> \> $mu_y = C A mu_x;$\\
\> \> $sigma_y = CA sigma_x A'C' + CQC' + R;$\\
\> \> $o[i] = $ {\tt gaussianNumberGeneration}$(mu_y, sigma_y);$\\
\> \> $mu_x = A mu_x$\\
\> \> $sigma_x = A  sigma_x A + Q$\\
\> {\tt end}\\
\> {\tt return} $o;$
\end{tabbing}
}
\caption{Generating Observations from an LDS Algorithm}
\vspace{0.6cm}
\end{figure}

>>>>

\begin{algorithm}[h]
\begin{pseudocode}
\codename $\code{lds\_inference}\left(Y,A,C,Q,R,x,P\right)$\\
\codeline \code{for } $t=1$ to $T$ ;forward recursion \\
\codeline \> $x_t^{t-1} = Ax_{t-1}^{t-1}$\\
\codeline \> $P_t^{t-1} = AP_{t-1}^{t-1}A' + Q \code{ if } t>1$\\
\codeline \> $K_t = P_t^{t-1}C'(CP_t^{t-1}C'+R)^{-1}$\\
\codeline \> $x_t^t = x_t^{t-1} + K_t(y_t - Cx_t^{t-1})$\\
\codeline \> $P_t^t = P_t^{t-1} - K_tCP_t^{t-1}$ \\
\codeline $\hat{P}_{T,T-1}=(I-K_TC)AP_{T-1}^{T-1}$ ;initialize \\
\codeline $\code{for } t=T$ to $2$ ; backward recursion \\
\codeline \> $J_{t-1} = P_{t-1}^{t-1}A'(P_t^{t-1})^{-1}$\\
\codeline \> $\hat{x}_{t-1} = x_{t-1}^{t-1} + J_{t-1}(\hat{x}_{t} - Ax_{t-1}^{t-1})$\\
\codeline \> $\hat{P}_{t-1} = P_{t-1}^{t-1} + J_{t-1}(\hat{P}_t - P_t^{t-1})J_{t-1}'$\\
\codeline \> $\hat{P}_{t,t-1} = P_t^t J_{t-1}' + J_t(\hat{P}_{t+1,t} - AV_t^t)J_{t-1}' \code{ if } t<T$\\
\codeline $\code{ return } \hat{x}_t, \hat{P}_t, \hat{P}_{t,t-1} \forall t$
\end{pseudocode}
\end{algorithm}

\begin{algorithm}[h]
\begin{pseudocode}
\codename $\code{lds\_learn}\left(Y,k,\epsilon\right)$\\
\codeline $A, C, Q, R,x_1^0, P_1^0$\\
\codeline $\alpha = \sum_t y_ty_t'$\\ 
\codeline $\code{while}$ change in likelihood $> \epsilon$\\ 
\codeline \> ; E-step \\
\codeline \> $\code{lds\_inference}(Y,A,C,Q,R,x_1^0,P_1^0)$ \\
\codeline \> $\delta = 0, \gamma = 0, \sigma = 0$ \\
\codeline \> $\code{for} t = 1$ to $T$\\ 
\codeline \> \> $\delta = \delta + y_t\hat{x}_t$\\ 
\codeline \> \> $\gamma = \gamma + \hat{x}_t\hat{x}_t' + \hat{P}_t$\\ 
\codeline \> \> $\beta = \beta + \hat{x}_t\hat{x}_{t-1}' + P_{t,t-1} \code{ if } t>1$\\ 
\codeline \> $\gamma_1 = \gamma - \hat{x}_T\hat{x}_T' -\hat{P}_T$\\ 
\codeline \> $\gamma_2 = \gamma - \hat{x}_1\hat{x}_1' - \hat{P}_1$\\ 
\codeline \> ; M-step\\ 
\codeline \> $C = \delta \gamma^{-1}$\\ 
\codeline \> $R = (\gamma - C\delta')/T$\\ 
\codeline \> $A = \beta\gamma_1^{-1}$\\ 
\codeline \> $Q = (\gamma_2 - A\beta') / (T-1)$\\ 
\codeline \> $x_1^0 = \hat{x}_1$\\ 
\codeline \> $P_1^0 = \hat{P}_1$\\ 
\codeline $\code{return } A, C, Q, R, x_1^0, P_1^0$
\end{pseudocode}
\end{algorithm}

@@

\begin{algorithm}[h]
\begin{pseudocode}
\codename $\code{observation\_generation}\left(A,C,R,Q,\mu,\Sigma,T,N\right)$\\
\codeline $\mu_x = \mu[T-1];$\\
\codeline $\sigma_x = \sigma[T-1];$\\
\codeline $\code{for } i = 1$ to $N$\\
\codeline \> $\mu_y = C A \mu_x;$\\
\codeline \> $\sigma_y = CA \sigma_x A'C' + CQC' + R;$\\
\codeline \> $o[i] = \code{gaussian\_number\_generation}(\mu_y, \sigma_y)$\\
\codeline \> $\mu_x = A \mu_x$\\
\codeline \> $\sigma_x = A  \sigma_x A + Q$\\
\codeline $\code{return } o$
\end{pseudocode}
\end{algorithm}
